---
title: "Mouse_modeling"
author: "LeZhang"
date: "6/13/2018"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: console
---

# Setup
This is a automatic report script. Users only need to place this Rmd to the same folder as you store mouse training data and knit. A html report will be generated.

You need to specify your go/nogo ratio below
```{r , echo = T}
goRatio = 0.50

```

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE, warning= F, message= F}
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F)
library(tidyverse)
library(magrittr)
library(stringr)
library(gridExtra)
library(broom)
```

# Data structure

## Mice from different days analyzed

```{r read, warning= F, }
files = list.files(pattern = "behavior.csv$")


mouseID  = as.factor(str_extract(files, "[^_]+"))

mouse_time = as.Date(gsub(".*_([0-9]+-[0-9]+-[0-9]+)_.*$", "\\1", files)) %>%
    tibble(date = .) %>%
    mutate(mouseID = mouseID) %>%
    group_by(mouseID) %>%
    mutate(day = rank(date))

dfs = lapply(files,
             function(x)
             read_csv(x,
             col_types = cols(
             X11 = col_skip(),
             lick_timestamps = col_character()
             )))
             
dfs = lapply(seq_along(dfs),
             function(x)
                 mutate(dfs[[x]], mouseID = mouseID[x]))

dfs  = lapply(seq_along(dfs),
              function(x)
                  mutate(dfs[[x]], day = mouse_time$day[x]))
             
df = as.tibble(plyr::rbind.fill(dfs))

print(mouse_time) 
```

## Structure of data set
```{r glimpse}
dim(df)
names(df)
```

## clean
Check structure again after removing `NA`
```{r clean1}
#replace all NA in injection to 0
df$InjectOrNot = 0

#temp change NA in time to 0
df$lick_timestamps[is.na(df$lick_timestamps)] = "99"

# remove last a few empty rows
df = drop_na(df)
dim(df)

# only leave the first lick time
df$lick_timestamps = as.numeric(gsub(pattern = "[ ].*", "\\1", df$lick_timestamps))

#add a lick no-lick column
df = add_column(df, lick_nolick = as.factor(ifelse(df$lick_timestamps != 99, "lick", "nolick")))

# 
```


```{r clean2}
# add category column


df = add_column(df, B.type = as.character(NA))


df  %<>%
    mutate(B.type = if_else(hit_trials == 1,
    "hit",
    if_else(
        miss_trials == 1,
        "mis",
        if_else(cr_trials == 1,
                "cr",
                if_else(fa_trials == 1,
                    "fa",
                    if_else(premature_trials == 1,
                        "premature",
                        "error")
                )
        )
    )
))
    
df %>% group_by(B.type, mouseID) %>% tally() 

#delete error rows
df = filter(df, B.type != "error")

```

# Basic stats of in plots 
```{r stats}
#check again with plot

df %>% 
  group_by(B.type, mouseID, day) %>%
  tally() %>% 
  group_by( mouseID, day) %>%
  mutate(percent = round(n/sum(n)*100, 0)) %T>%
            {print(
              ggplot(.)+
                geom_bar(aes(x = B.type, y = n, fill = B.type), stat = "identity") +
                geom_text(aes(x = B.type, y = n, label = percent)) + 
                facet_grid(mouseID ~ day) + 
                theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +
                ggtitle("type counts and percent of different mice in different days")
                )
             print(
              ggplot(.)+
                geom_line(aes(x = day, y = percent, color = B.type)) +
                #geom_text(aes(x = B.type, y = n, label = percent)) + 
                facet_grid(.~ mouseID) +
                scale_x_continuous(breaks=1:max(mouse_time$day)) + 
                ggtitle("type trend in percent of different mice in different days")
                )
            } 



```
> Add one column: correct.b - correct behavior, correct 1, fail 0

> Add one column: correct.b.ma - correct behavior and premature, correct 1, fail -1, premature 0



```{r}
# add one column: correct.b
# correct behavior
df %<>% 
  mutate(correct.b = if_else(condition = hit_trials == 1,
                             true      = 1,
                             false     = if_else(cr_trials == 1, 1, 0)
                             )
         ) 
## premature count as failure here, but need to be removed when build model

# add one column: correct.b.ma 

df %<>% 
  mutate(correct.b.ma = if_else(condition = hit_trials == 1,
                             true      = 1,
                             false     = if_else(cr_trials == 1, 1,
                                                 if_else(premature_trials == 1,
                                                         0, -1))
                             )
         )
```


# Go : NoGo Ratio & Correct Ratio
```{r ,message = F}
df %>% filter(premature_trials != 1) %>%
  ggplot() +
    geom_bar(aes(x = factor(go_trials),
                 y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..],
                 fill = factor(go_trials)
                 ),
             position = "stack"
             ) +
    facet_grid(mouseID ~ day) + 
    xlab("Nogo vs. GO") +
    ylab("percent") + 
    ggtitle("Go and NoGO of different mice in different days")

df %>% 
  filter(premature_trials != 1) %>%
  group_by(day, mouseID) %>%
  summarise(ratio = round(sum(go_trials)/length(go_trials), 2),
            correct.b.rate = round(sum(correct.b)/length(correct.b), 2)
            ) %>%
  add_column(sample = seq(nrow(.))) %T>%
    {
      p1 = ggplot(.) +
        geom_text(aes(x = sample, y = ratio, label = ratio, color = mouseID), size = 3) +
        geom_smooth(aes(x = sample, y = ratio, color = mouseID), alpha = 0.05) +
        ylim(0.4, 0.6) +
        ggtitle("Go : NoGO trend")
           
      
      p2 = ggplot(.) +
        geom_boxplot(aes(x = mouseID, y = ratio, color = mouseID)) +
        geom_point(aes(x = mouseID, y = ratio, color = mouseID))+
        ylim(0.4, 0.6) +
        ggtitle("Go : NoGO boxplot")
      
      p3 = ggplot(.) +
        geom_point(aes(x = day, y = correct.b.rate, color = mouseID))+
        geom_smooth(aes(x = day, y = correct.b.rate, color = mouseID), alpha = 0.01)+
        ylim(0.5, 0.9) +
        ggtitle("Correct rate by mouse and day")
      
    grid.arrange(p1, p2, p3, nrow = 3)
    } %>%
  ungroup() %>%
  group_by(mouseID) %>%
  do(tidy(t.test(.$ratio, data=.)))

  
```


## 95% interval of Go: NoGo = `r goRatio`

based on 10 times max days( `r max(mouse_time$day)` days) random sampling with 350 trials per day, the 95 confidence intervel is as below:
```{r , message=F}


replicate(max(mouse_time$day) * 10,
          mean(rbinom(350, 1, goRatio))) %>%
          t.test() %>%
          tidy()
```


# Modeling
All models tested with 25 fold bootstrap validations.



```{r}
library(caret)

controls <- trainControl(method = "cv", number = 5)
```

## Fit data to a logistic model
While buiding the model, prematures are removed, the previous go/nogo number is treated as previous trial. 
```{r simple logistic}
logistic_simple = df %>%
  filter(premature_trials != 1) %>%
  select(mouseID, day, TrialNum, correct.b, correct.b.ma)

logistic_simple[,paste0("previous.", 1:10, "t")] = NA

  
logistic_simple_list = logistic_simple %>%
  group_by(mouseID, day) %>%
  plyr::dlply(., .variables = c("mouseID", "day"), identity)

##
i = NULL
j = NULL
nList = NULL
for (nList in seq_along(logistic_simple_list)){
  for (j in seq_len(nrow(logistic_simple_list[[nList]]))){
    for (i in seq_len(10)) {
      if (j - i > 0.5){
        logistic_simple_list[[nList]][j, 5+i] = logistic_simple_list[[nList]][j-i, 4]
      }
    }
  }
}
#check
#logistic_simple_list[[1]] 

#
logistic_summary = matrix(rep(NA,10*length(logistic_simple_list)),
                             ncol = length(logistic_simple_list)) %>%
                   as.tibble()

names(logistic_summary) = names(logistic_simple_list)


for (nList in seq_along(logistic_simple_list)){
  logistic_simple_list[[nList]] %>%
  as.tibble() %>%
  mutate_all(., .funs = factor) %>%
  drop_na() %>%
  select(- mouseID, - day, - TrialNum, - correct.b.ma) ->
  logistic_simple_df
  

  sapply(2:11,function(i){
      logistic_simple_df_sub = select(logistic_simple_df,1:i)
      train(correct.b ~ .,
            data=logistic_simple_df_sub,
            method="glm",
            family = binomial,
            na.action = na.pass
           ) %>% 
     .[["results"]] %>%
     .[2] 
  }) %>%
  unlist() ->
  logistic_summary[,nList]
  
}

#
logistic_summary %<>%
  mutate(trace_back_num = paste0("back.", str_pad(1:10, 2, pad = "0")))

#
logistic_summary  %>%
  t() %>%
  as.tibble() ->
  logistic_summary_t
  
names(logistic_summary_t) = logistic_summary_t[nrow(logistic_summary_t),]
logistic_summary_t = logistic_summary_t[-nrow(logistic_summary_t),]

logistic_summary_t_plot = 
  gather(logistic_summary_t) %>%
  mutate(key = as.factor(key), value = as.numeric(value)) %>%
  mutate(type = "original")


#
logistic_summary_plot = 
logistic_summary %>%
  select(-trace_back_num) %>%
  gather() %>%
  {str_split(.$key, "\\.", simplify = T)} %>%
  cbind(.,
        logistic_summary %>%
          select(-trace_back_num) %>%
          gather())
  
names(logistic_summary_plot) = c("mouseID", "day", "key", "value")

logistic_summary_plot %<>% select( -key)

logistic_summary_plot %<>%
  mutate(trial = rep(1:10, nrow(logistic_summary_plot) / 10)) %>%
  mutate(mouseID = as.factor(mouseID),
         day     = as.factor(day),
         trial   = as.factor(trial),
         value   = as.numeric(value),
         type    = "original")
```

# Let's do the shuffle before ploting

Shuffle 5 times and take the mean

Code maybe hiding in final report
```{r}
#shuffle
shuf_logistic = df %>%
  filter(premature_trials != 1) %>%
  select(mouseID, day, TrialNum, correct.b, correct.b.ma)

shuf_logistic[,paste0("previous.", 1:10, "t")] = NA

Shuffle = function(shuf_logistic){
shuf_logistic %<>%
  group_by(mouseID, day) %>%
  mutate(correct.b = sample(correct.b))

  
shuf_logistic_list = shuf_logistic %>%
  group_by(mouseID, day) %>%
  plyr::dlply(., .variables = c("mouseID", "day"), identity)

##
i = NULL
j = NULL
nList = NULL
for (nList in seq_along(shuf_logistic_list)){
  for (j in seq_len(nrow(shuf_logistic_list[[nList]]))){
    for (i in seq_len(10)) {
      if (j - i > 0.5){
        shuf_logistic_list[[nList]][j, 5+i] = shuf_logistic_list[[nList]][j -i, 4]
      }
    }
  }
}

#
logistic_shuf_summary = matrix(rep(NA,10*length(shuf_logistic_list)),
                             ncol = length(shuf_logistic_list)) %>%
                   as.tibble()

names(logistic_shuf_summary) = names(shuf_logistic_list)


for (nList in seq_along(shuf_logistic_list)){
  shuf_logistic_list[[nList]] %>%
  as.tibble() %>%
  mutate_all(., .funs = factor) %>%
  drop_na() %>%
  select(- mouseID, - day, - TrialNum, - correct.b.ma) ->
  logistic_simple_df
  

  sapply(2:11,function(i){
      logistic_simple_df_sub = select(logistic_simple_df,1:i)
      train(correct.b ~ .,
            data=logistic_simple_df_sub,
            method="glm",
            family = binomial,
            na.action = na.pass
           ) %>% 
     .[["results"]] %>%
     .[2] 
  }) %>%
  unlist() ->
  logistic_shuf_summary[,nList]
  }
return(logistic_shuf_summary)
}

logistic_shuf_summary = replicate(5, list(Shuffle(shuf_logistic)))
    


logistic_shuf_summary = as.tibble(Reduce("+", logistic_shuf_summary)/length(logistic_shuf_summary))

logistic_shuf_summary %<>%
  mutate(trace_back_num = paste0("back.", str_pad(1:10, 2, pad = "0")))

logistic_shuf_summary  %>%
  t() %>%
  as.tibble() ->
  logistic_shuf_summary_t
  
names(logistic_shuf_summary_t) = logistic_shuf_summary_t[nrow(logistic_shuf_summary_t),]
logistic_shuf_summary_t = logistic_shuf_summary_t[-nrow(logistic_shuf_summary_t),]

logistic_shuf_summary_t_plot = 
  gather(logistic_shuf_summary_t) %>%
  mutate(key = as.factor(key), value = as.numeric(value)) %>%
  mutate(type = "shuffle")

#
logistic_shuf_plot = 
logistic_shuf_summary %>%
  select(-trace_back_num) %>%
  gather() %>%
  {str_split(.$key, "\\.", simplify = T)} %>%
  cbind(.,
        logistic_shuf_summary %>%
          select(-trace_back_num) %>%
          gather())
  
names(logistic_shuf_plot) = c("mouseID", "day", "key", "value")

logistic_shuf_plot %<>% select( -key)

logistic_shuf_plot %<>%
  mutate(trial = rep(1:10, nrow(logistic_shuf_plot) / 10)) %>%
  mutate(mouseID = as.factor(mouseID),
         day     = as.factor(day),
         trial   = as.factor(trial),
         value   = as.numeric(value),
         type    = "shuffle")
```


```{r , eval = F}
#get random and predictable sequence
Pred = function(percent){
df_pred = tibble(correct.b = rep(c(0, 1), 2000))

df_pred[,paste0("previous.", 1:10, "t")] = NA

df_pred %<>%
add_column(TrialNum = seq_len(nrow(df_pred)), .before = "correct.b")

# muate some sequence

df_pred$correct.b[(sample(4000) <= 40*percent)] = sample(df_pred$correct.b[(sample(4000) <= 400)])

i = NULL
j = NULL

  for (j in seq_len(nrow(df_pred))){
    for (i in seq_len(10)) {
      if (j - i > 0.5){
        df_pred[df_pred$TrialNum == j, 2+i] = df_pred[df_pred$TrialNum == j -i, 2]
      }
    }
  }



pred_summary = matrix(rep(NA,(length(df_pred) - 2)),
                             ncol = length(df_pred)-2) %>%
                   as.tibble()

names(pred_summary) = names(df_pred)[c(-1, -2)]


sapply(3:12,function(i){
    df_pred_sub = select(df_pred,2:12) %>%
      mutate_all(., .funs = factor) %>%
      drop_na()
    train(correct.b ~ .,
          data=df_pred_sub,
          method="glm",
          family = binomial,
          na.action = na.pass
         ) %>% 
   .[["results"]] %>%
   .[2]
  }
) %>%
  unlist() %>%
  return()
}

PredictionRate = map(seq(from = 10, to = 90, by = 10)
           ,Pred) %>%
  unlist %>%
  tibble
PredictionRate %<>%
  rename(rate    = '.') %>%
  mutate(percent = rep(seq(from = 10,
                           to   = 90,
                           by   = 10),
                       each = 10),
         history = rep(seq(1:10),
                       9)
         )
write.csv(PredictionRate, "PredictionRate.csv", row.names = F)
```


```{r}
PredictionRate = read_csv("PredictionRate.csv")

ggplot(PredictionRate)+
  geom_smooth(aes(x = history, y = rate, color = percent, group = percent)) +
  ggtitle("Relation between prediction rate vs. previous history and random percentage")
```


```{r}
#combine real and shuffle data set
logistic_plot = as.tibble(bind_rows(logistic_summary_plot, logistic_shuf_plot))

# train a random sequence
set.seed(1)
sequence  = rbinom(35000, 1, 0.5)
set.seed(2)
results = rbinom(35000, 1, 0.5)

```


```{r}
ggplot(logistic_summary_t_plot) +
  geom_boxplot(aes(x = key, y = value, group = key)) +
  geom_point(aes(x = key, y = value, group = key), color = "#4286f4")+
  geom_smooth(aes(x = key, y = value, group=1), alpha = 0.2)+
  geom_hline(aes(yintercept = mean(value), linetype = "mean"), color ="blue")+
  geom_hline(aes(yintercept = mean(logistic_shuf_summary_t_plot$value), linetype = "shuffle mean"), color = "red")+
  geom_hline(aes(yintercept = 0.5, linetype = "random"), color = "black")+
  scale_linetype_manual(name  = "line type", values = c(2, 2, 2), 
                        guide = guide_legend(override.aes = list(color = c("blue", "black", "red"))))





ggplot(logistic_plot)+
  geom_point(aes( x = trial, y = value, group = type, color = type))+
  geom_smooth(aes(x = trial, y = value, group = type, color = type), alpha = 0.2) +
  facet_grid(mouseID~day) 
  

```

## Fit data to a KNN model 

```{r knn}
df_knn = df %>%
  select(mouseID, day, TrialNum, correct.b.ma)

df_knn_shuf = df_knn
set.seed(1)
df_knn_shuf$correct.b.ma = sample(df_knn_shuf$correct.b.ma)

Pred_knn = function(df_knn){
  
  df_knn[,paste0("previous.", 1:10, "t")] = NA

  
knn_list = df_knn %>%
  group_by(mouseID, day) %>%
  plyr::dlply(., .variables = c("mouseID", "day"), identity)
##
i = NULL
j = NULL
nList = NULL
for (nList in seq_along(knn_list)){
  for (j in seq_len(nrow(knn_list[[nList]]))){
    for (i in seq_len(10)) {
      if (j - i > 0.5){
        knn_list[[nList]][knn_list[[nList]]$TrialNum == j, 4+i] = knn_list[[nList]][knn_list[[nList]]$TrialNum == j -i, 4]
      }
    }
  }
}

knn_summary = matrix(rep(NA,10*length(knn_list)),
                             ncol = length(knn_list)) %>%
                   as.tibble()

names(knn_summary) = names(knn_list)

for (nList in seq_along(knn_list)){
  knn_list[[nList]] %>%
  as.tibble() %>%
  mutate_all(., .funs = factor) %>%
  drop_na() %>%
  select(- mouseID, - day, - TrialNum) ->
  knn_simple_df
  
  sapply(2:11,function(i){
      knn_simple_df_sub = select(knn_simple_df,1:i)
      train(correct.b.ma ~ .,
            data=knn_simple_df_sub,
            method="knn",
            na.action = na.pass
            ) %>% {
              best = as.numeric(.[[6]])
              .[[4]] %>%
                filter(k == best) %>%
                .[1,2]
           }
  }) %>%
  unlist() ->
  knn_summary[,nList]
}


knn_summary %<>%
  mutate(trace_back_num = paste0("back.", str_pad(1:10, 2, pad = "0")))


knn_summary  %>%
  t() %>%
  as.tibble() ->
  knn_summary_t
  
names(knn_summary_t) = knn_summary_t[nrow(knn_summary_t),]
knn_summary_t = knn_summary_t[-nrow(knn_summary_t),]

knn_summary_t_plot = 
  gather(knn_summary_t) %>%
  mutate(key = as.factor(key), value = as.numeric(value))

p1 = ggplot(knn_summary_t_plot) +
  geom_boxplot(aes(x = key, y = value, group = key)) +
  geom_point(aes(x = key, y = value, group = key), color = "#4286f4")+
  geom_smooth(aes(x = key, y = value, group=1), alpha = 0.2)+
  geom_hline(aes(yintercept = mean(value)))

knn_summary_plot = 
knn_summary %>%
  select(-trace_back_num) %>%
  gather() %>%
  {str_split(.$key, "\\.", simplify = T)} %>%
  cbind(.,
        knn_summary %>%
          select(-trace_back_num) %>%
          gather())
  
names(knn_summary_plot) = c("mouseID", "day", "key", "value")

knn_summary_plot %<>% select( -key)

knn_summary_plot %<>%
  mutate(trial = rep(1:10, nrow(knn_summary_plot) / 10)) %>%
  mutate(mouseID = as.factor(mouseID),
         day = as.factor(day),
         trial = as.factor(trial),
         value = as.numeric(value))


p2 = ggplot(knn_summary_plot)+
  geom_point(aes(x = trial, y = value), color = "#4286f4")+
  geom_smooth(aes(x = trial, y = value, group = 1), alpha = 0.2) +
  facet_grid(mouseID~day)

print(p1)
print(p2)
}

Pred_knn(df_knn_shuf)
```
Fit Knn model with correct, not correct and pre-mature
```{r}
Pred_knn(df_knn)
```

```{r complex logistic, eval= F}

track1_model = df %>% 
  select(-B.type, -correct.b.ma) %>% 
  filter(TrialNum %in% c(1:3)) %>% 
  mutate(correct.b = replace(correct.b, TrialNum %in% c(1, 2), NA)) %>%
  group_by(mouseID,day) %>%
  fill(correct.b, .direction = "up") %>%
  do(modelfit = train(factor(correct.b) ~ . - mouseID - day - TrialNum,
                      data=., method="glm",
                      family = binomial,
                      trControl=trainControl(method="none")
                          )
     )
tidy(track1_model, modelfit)
track1_model$modelfit[1:10]
```

## correlation
corplot of each variable used agianst 'if correct behavior' as the y variable.

In the following plot, `correct.b.ma` is the response variable.

```{r }
ColorCorplot = colorRampPalette(c("#67001F", "#B2182B", "#D6604D",
                                  "#92C5DE",
                                  "#4393C3", "#2166AC", "#053061"))


df %>% 
  select(go_trials,
         lick_timestamps,
         day,
         correct.b.ma) %>%
  cor() %>%
  corrplot::corrplot(.,
                     method="number",   
                     type="upper",
                     col = ColorCorplot(100)
                     )

```

## Autocorrelation

```{r}
library(tseries)
df %>%
  filter(premature_trials != 1) %>%
  group_by(mouseID,day) %>%
  count() %>%
  nrow() ->
  nsamples

par(mfrow=c(5,ceiling(nsamples/5)))

df %>%
  filter(premature_trials != 1) %>%
  group_by(mouseID,day) %>%
  select(correct.b) %>%
  do(., pacf(.$correct.b, lag.max = 10, plot = T))
dev.off()

df %>%
  filter(premature_trials != 1) %>%
  is.na(.)
pacf(df$correct.b, plot = F, lag.max = 30) %>% plot
```



# Ignore below

## Fit a linear model of `correct.b.ma`
```{r , eval= F}


lm_model = df %>% 
              select(go_trials,
                     lick_timestamps,
                     day,
                     correct.b.ma) %>%
              train(correct.b.ma ~ ., data=., method="glm", trControl = controls)

lm_model
summary(lm_model)

```

## A logistic model classification with `correct.b`
The max outcome from glm is 2: 0 & 1; so can't use `correct.b.ma`. Prematures are removed

```{r , eval= F}
logit_model = df %>% 
                select(go_trials,
                       lick_timestamps,
                       day,
                       premature_trials,
                       correct.b) %>%
                filter(premature_trials == 0) %>%
                select(-premature_trials) %>%
                train(factor(correct.b) ~ .,
                      data=., method="glm",
                      family = binomial(),
                      trControl = controls)
logit_model
summary(logit_model)      
```

## KNN model with `B.type`
Use KNN to fit b.type which can handle more than 2 types.

```{r , eval= F}
knn_model = df %>% 
              select(go_trials,
                     lick_timestamps,
                     day,
                     B.type) %>%
              train(factor(B.type) ~ ., data=., method="knn", trControl = controls)

knn_model

ggplot(df)+
  geom_point(aes(x = B.type, y = lick_timestamps))
```

